---
title: "Phylogenetic Analysis"
author: "Arie Fridrich"
format: html
editor: visual
---

## Introduction

I made this workflow to help anyone interested in having a practical guide for performing phylogenetic analysis at a relatively large scale and know which factors to consider. I wish to credit Dr. Chung Hyun Cho for his patience and teaching the conceptss presented in this workflow.

## Invest in the collection of data sets

Download the the [eukaryota_odb10.2020-09-10.tar.gz](https://busco-data.ezlab.org/v5/data/lineages/eukaryota_odb10.2020-09-10.tar.gz) data base (or any other of interest) from <https://busco.ezlab.org/>.

```{bash}
cd /groups/berger/user/arie.fridrich/Paper_MpAtDDM1/ddm1evo/Phylo/
wget https://busco-data.ezlab.org/v5/data/lineages/eukaryota_odb10.2020-09-10.tar.gz
```

This dataset contains sequences of a set of conserved eukaryotic genes that can be used to asses completeness of genomes. Now we need to put or hand on a set of fasta files for the analysis. I will look at Nematostella, Human, Arabidopsis, Marchantia, Physco and Cm.

Proteomes should have represetative gene models and ideally not isoforms.

```{bash}
#Nematostella: NV2 versio from stowers. 
#from eukProt V3. Human: EP00075, Nematostella: EP00110.
wget http://evocellbio.com/SAGdb/images/TCS.tar.gz
#Arabidopsis: https://www.arabidopsis.org/download/index-auto.jsp?dir=%2Fdownload_files%2FProteins%2FAraport11_protein_lists/Araport11_pep_20220914.gz
```

Example how I parse a fasta file to remove isoforms in R:

```{r message = FALSE, warning = FALSE}
# I will be using the tidysq package for importing the fasta file to R https://rdrr.io/github/michbur/tidysq/, so first I will install it:
install.packages("tidysq")

# Load packeges need for the parsing:
library(tidyverse)
library(tidysq)


setwd("/groups/berger/user/arie.fridrich/Paper_MpAtDDM1/ddm1evo/Phylo/")


```

The Purpose ssd fRunning Cosde

When you click the **Render** button a document will be generated that includes both content and the output of embedded code. You can embed code like this:

```{r}
1 + 1
```

You can add options to executable code like this

```{r}
#| echo: false
2 * 2
```

The `echo: false` option disables the printing of code (only output is displayed).
